Generality in food webs scales with species richness, not latitude

### 2015-01-29 manuscript/latitudepaper.tex
using greyscale fig for printoff

We then assessed
the impact of latitude on this relationship; specifically, the effect of latitude on the scaling exponent $\beta$. 
In addition to latitude, we included a factorial variable for ecosystem type (stream, lake, terrestrial, marine, or estuary) to account for recognized variation in food web structure across ecosystem types \citep{}.
% We also considered other trends in the data that could affect relationships between latitude and species richness.
% To account for changing standards in data collection over time \citep{Dunne2006} we included an effect of year of
% publication. As 11 of our XX webs included humans, and therefore might have different properties than other networks,
% we included a factorial effect of the presence of humans. 
Note that some of our webs describe the same area in different
seasons or different years. [[removed the site-level random effect because I don't want it to steal latitude information.]]

@@

@@

\caption{Standard effects and their $p$-values for the best-fitting models of links per species ($Z$), generality ($G$), and vulnerability ($V$).}
\label{Bestfits}
\begin{tabular}{l | l l | l l | l l }
\multirow{2}{*}{Effect} & \multicolumn{2}{c}{$Z$} & \multicolumn{2}{c}{$G$} & \multicolumn{2}{c}{$V$}\\
                & $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value & \\
\hline
Intercept               & -0.561 & \textless0.001 & -0.299 & \textless0.001 & -0.561 & \textless0.001 \\
log(species)            &  0.733 & \textless0.001 &  0.622 & \textless0.001 &  0.733 & \textless0.001 \\
log(species):Lake       & -0.118 & 0.047          & -0.115 & 0.078          & -0.118 & 0.047 \\
log(species):Marine     &  0.051 & 0.040          &  0.120 & \textless0.001 &  0.051 & 0.040 \\
log(species):Stream     & -0.161 & 0.014          & -0.057 & 0.421          & -0.161 & 0.014 \\
log(species):Latitude   & -0.001 & 0.081          & -0.001 & 0.273          & -0.001 & 0.081 \\
log(species):Lat:Lake   &  0.006 & \textless0.001 &  0.006 & \textless0.001 &  0.006 & \textless0.001 \\
log(species):Lat:Stream &  0.004 & 0.018          &  0.003 & 0.059          &  0.004 & 0.018 \\
\hline
$R^{2} (adjusted)$   & \multicolumn{2}{c}{0.706} & \multicolumn{2}{c}{0.658} & \multicolumn{2}{c}{0.706} \\
\end{tabular}
\end{table}

[[This'll need to be redone]]
% The tendency of food-web structure to exhibit scaling relationships with both
% species richness and connectance has been well-established \citep{}. As
% species richness in particular is also known to vary systematically over
% latitude \citep{}, intuitively one might suspect that any relationship
% between food-web properties such as generality might be due to the latitudinal
% gradient in species richness. Here, we found that while there were latitudinal
% gradients in species richness and generality, latitude was not related to
% either the scaling relationship between generality and species richness or the
% residuals of this scaling relationship.

% The latitudinal gradient in species richness in this data set ran in the opposite 
% direction to the general trends for (list types of taxa). This may be due to the 
% fact that latitudinal gradient in species richness are usually studied within a
% taxonomic or trophic group (e.g., birds, vascular plants), whereas food webs 
% measure many taxa concurrently. It is possible that while many taxa generally increase
% towards the tropics, the number of species across all taxa that co-occur within an 
% study site decreases. Alternatively, it is possible that by using trophic-species versions
% of our food webs, we have artificially reversed the species-richness gradient. [[re-run analyses
% with species-species webs]]. This would imply that, while there may be more species in the tropics,
% many of those species tend to have identical predator and prey sets to co-occuring species. In either
% case, it does not seem likely that the increase in
% generality with latitude that we observed is due to an underlying tendency for
% species to pursue a wider variety of prey at higher latitudes. Rather, the
% trend in generality can be more parsimoniously explained by the trend in
% species richness.

% Our findings here add further doubt to the niche-breadth hypothesis, as
% narrower niches in the tropics were associated with lower species richness in
% this data set. It appears instead that, as more species are available, each
% species tends to consume more prey, be consumed by more predators, and have
% more links in general (check with figs). This makes intuitive sense as a more
% species-rich community is likely to have more species with similar behaviour
% and morphology that thereby have overlapping predator or prey sets.

% Parallel increases in generality and species richness across latitude have
% also been noted in studies of plant-pollinator networks (although in that case
% species richness was highest in the tropics) \citep{Schleuning2012}. Although
% the authors of that study do not explicitly examine the relationship between
% specialization and network size, they do cite increasing plant species
% richness as a potential driver for increasing generalization
% \citep{Schleuning2012}. Although the scaling of mutualist network structure
% does not appear to have been explored as systematically as that of unipartite
% food webs, some poperties such as nestedness \citep{Vazquez2004} and XX
% \citep{} have been shown to scale with species richness. As such scaling
% appears to be common to both antagonistic and mutualistic food webs, and
% trends in species richness can explain counterintuitive trends in network
% structure \citep{Schleuning2012}, it seems likely that many latitudinal trends
% in network structure are at least partly due to trends in species richness.

% %Nestedness increases with A+P, increases number of extreme specialists AND number of generalists. Not clear how degree should scale with S according to scaling with nestedness.

% \section*{Conclusion}

% It was not our initial intention to question the validity of the wide range of
% studies exploring variation in network structure across changing environments \citep{}.
% However, if those studies have not taken into account changes in species
% richness over environmental gradients, it is possible that observed changes in
% food-web structure are due only indirectly to the environmental variables of
% interest. Effects of environmental variables on network structure should
% therefore be interpreted with caution unless effects of changing species
% richness are made explicit.

@@

\begin{center}
\begin{table}[!h]
\caption{Symbols for and descriptions of food-web properties.}
\label{symboltable}
\begin{tabular}{c l l}
\hline
Symbol      & Name & Description \\
\hline
S & Species richness & Number of trophic species \\
LS & Links per species & Mean number of links per taxon \\
Gen & Generality & Mean number of resources per taxon \\
Vul & Vulnerability & Mean number of consumers per taxon \\
\hline
\end{tabular}
\end{table}
\end{center}
% NODF & Average Nestedness & Degree to which interactions among specialists are subsets or interactions among generalists \\
% Chain & Chain length & Mean food chain length \\
% Cluster & Clustering coefficient & Mean clustering coefficient (probability that two taxa linked to the same taxon are also linked) \\

\begin{center}
\begin{table}[!h]
\caption{Full models for S, LS, and C.}
\label{symbols}
\begin{tabular}{c c c c c c c}
\hline
 & \multicolumn{2}{c}{S} & \multicolumn{2}{c}{C} & \multicolumn{2}{c}{LS} \\
\hline
Effect & Estimate & $p$-value & Estimate & $p$-value & Estimate & $p$-value \\
\hline
Latitude 			& 0.002  & 0.873  & -0.006 & 0.480  & -0.004 & 0.613 \\ 
Estuary 			& 1.078  & 0.022  & -0.809 & 0.012  & 0.269 & 0.362 \\
Freshwater 			& 1.347  & \textless0.001  & -1.229 & \textless0.001 & 0.118 & 0.394 \\
Marine 				& 1.127  & \textless0.001 & -0.761 & \textless0.001 & 0.366 & 0.012 \\
Terrestrial 		& 1.097  & \textless0.001 & -0.919 & \textless0.001 & 0.178 & 0.354\\
Latitude:Freshwater & 0.002  & 0.902  & 0.010  & 0.299  & 0.012 & 0.194\\
Latitude:Marine 	& 0.003  & 0.847  & 0.001  & 0.902  & 0.004 & 0.662 \\
Latitude:Terrestrial & 0.001 & 0.923  & 0.003  & 0.724  & 0.005 & 0.597 \\
\end{tabular}
\end{table}
\end{center}

% \newpage

% \section*{Figures}

% \begin{figure}[!h]
% \label{gen_all}
% \includegraphics[width=.9\textwidth]{Figures/gen_all.eps}
% \caption{Relationships between mean generality and latitude for food webs with moderate connectance (0.1, mean of empirical webs) at high (S=109, solid line), moderate (S=36, dashed line), and low (S=5, dotted line) species richness.}
% \end{figure}

% \begin{figure}[!h]
% \label{TL_all}
% \includegraphics[width=.9\textwidth]{Figures/TL_all.eps}
% \caption{Relationships between mean trophic level and latitude for food webs with moderate connectance (0.1, mean of empirical webs) at high (S=109, solid line), moderate (S=36, dashed line), and low (S=5, dotted line) species richness.}
% \end{figure}

>>>>

We then assessed
the impact of latitude on this relationship; specifically, the effect of latitude on the scaling exponent $\beta$. 
In addition to latitude, we included a factorial variable for ecosystem type (stream, lake, terrestrial, marine, or estuary) to account for recognized variation in food web structure across ecosystem types \citep{}.
% We also considered other trends in the data that could affect relationships between latitude and species richness.
% To account for changing standards in data collection over time \citep{Dunne2006} we included an effect of year of
% publication. As 11 of our XX webs included humans, and therefore might have different properties than other networks,
% we included a factorial effect of the presence of humans. 
Note that some of our webs describe the same area in different
seasons or different years. 
% [[removed the site-level random effect because I don't want it to steal latitude information.]]

@@

The tendency of food-web structure to exhibit scaling relationships with both
species richness and connectance has been well-established \citep{}. As
species richness in particular is also known to vary systematically over
latitude \citep{}, intuitively one might suspect that any relationship
between food-web properties such as generality might be due to the latitudinal
gradient in species richness. In this dataset, however, we did not find
overall latitudinal gradients in species richness, links per species, 
generality, or vulnerability. This fact notwithstanding, latitude did affect the
scaling relationships between species richness and the other network properties.

The lack of latitudinal gradients in food-web properties in this dataset contrasts
strongly with other studies which have found consistent variation in species richness,
[[other properties]] \citep{}. As numbers of species and links included in a food web
vary strongly with sampling effort as well as with the underlying diversity of the study
area, it is possible that the lack of latitudinal trends here is a result of researchers
tending to expend similar amounts of sampling effort across studies. This could result in
food webs describing species-rich tropical communities omitting more species and links
than species-poor arctic communities. While unusual, the lack of association between species richness
and latitude in any ecosystem type nevertheless means that any effect of latitude on scaling relationships
between species richness and other properties is not being driven by underlying variation in
species richness, allowing us to more clearly assess effects of latitude on scaling.

Our scaling results provide only equivocal support for the niche-breadth hypothesis.
In particular, scaling relationships between species richness and links per species,
generality, or vulnerability strengthened with increasing latitude for lake and stream
food webs but did not vary significantly with latitude in estuarine, terrestrial, and
marine communities. In lakes and streams, the niche hypothesis' proposal that niches
are narrower in the tropics fits the observed scaling patterns. The increase in links
per species as food-web size increases is much slower in low-latitude communities than
in their high-latitude counterparts. From this, we can infer a higher proportion of
specialists with few predators in tropical communities. This echos earlier research [[details]]
\citep{}. In contrast, temperate and polar communities appear to contain more generalist
species. As food webs are generally aggregated over space and time, the stronger
increase in links per species with species richness could be driven by species switching
between more prey species that are only intermittently available in areas with highly
variable climates.

\section*{Conclusion}

[[something catchy to sum up]]

\begin{center}
\begin{table}[!h]
\caption{Symbols for and descriptions of food-web properties.}
\label{symboltable}
\begin{tabular}{c l l}
\hline
Symbol      & Name & Description \\
\hline
$S$ & Species richness & Number of trophic species \\
$Z$ & Links per species & Mean number of links per taxon \\
$G$ & Generality & Mean number of resources per taxon \\
$V$ & Vulnerability & Mean number of consumers per taxon \\
\hline
\end{tabular}
\end{table}
\end{center}

@@

\caption{Standard effects and their $p$-values for the best-fitting models of links per species ($Z$), generality ($G$), and vulnerability ($V$).}
\label{Bestfits}
\begin{tabular}{l | l l | l l | l l }
Effect          & \multicolumn{2}{c|}{$Z$} & \multicolumn{2}{c|}{$G$} & \multicolumn{2}{c}{$V$}\\
                & $\beta$ & $p$-value     &  $\beta$ & $p$-value    &  $\beta$ & $p$-value  \\
\hline
Intercept               & -0.561 & \textless0.001 & -0.299 & \textless0.001 & -0.561 & \textless0.001 \\
log(species)            &  0.733 & \textless0.001 &  0.622 & \textless0.001 &  0.733 & \textless0.001 \\
log(species):Lake       & -0.118 & 0.047          & -0.115 & 0.078          & -0.118 & 0.047 \\
log(species):Marine     &  0.051 & 0.040          &  0.120 & \textless0.001 &  0.051 & 0.040 \\
log(species):Stream     & -0.161 & 0.014          & -0.057 & 0.421          & -0.161 & 0.014 \\
log(species):Latitude   & -0.001 & 0.081          & -0.001 & 0.273          & -0.001 & 0.081 \\
log(species):Lat:Lake   &  0.006 & \textless0.001 &  0.006 & \textless0.001 &  0.006 & \textless0.001 \\
log(species):Lat:Stream &  0.004 & 0.018          &  0.003 & 0.059          &  0.004 & 0.018 \\
\hline
$R^{2} (adjusted)$   & \multicolumn{2}{c|}{0.706} & \multicolumn{2}{c|}{0.658} & \multicolumn{2}{c}{0.706} \\
\end{tabular}
\end{table}

@@

\section*{Figures}

\begin{figure}[!h]
\label{Latfigs}
\includegraphics[width=.9\textwidth]{Figures/properties_vs_lat.eps}
\caption{Relationships between latitude and species richness, mean links per species, mean generality, and mean vulnerability. No food web property showed a significant latitudinal gradient in this dataset (N=259 food webs).}
\end{figure}

\begin{figure}[!h]
\label{scalingfig}
\includegraphics[width=.9\textwidth]{Figures/scaling_with_S_grey.eps}
\caption{Relationships between species richness and mean links per species, mean generality, and mean vulnerability (N=259 food webs). In estuarine, terrestrial, and marine communities the scaling relationship with species-richness did not vary significantly across latitude for any food-web property. In lake and stream communities, each food-web property showed stronger scaling with species richness as latitude increased (N=259 food webs). }
\end{figure}
